### 2023-11-27 analysis/blots.qmd
adjust plot and bar themes

---
title: "Protein Expression"
format: 
  html:
    embed-resources: true 
    echo: false
    warning: false
---

```{r}
#| label: setup
#| include: false

library(here)
library(targets)
library(tidyverse)
devtools::load_all()

i_am("analysis/blots.qmd")

tar_config_set(store = here("_targets"))
```

```{r}
#| label: test
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_dual_col1a1) + 
  theme(legend.position = "bottom")
```