### 2023-11-27 analysis/blots.qmd
add blot graphs to analysis qmd

---
title: "Protein Expression"
format: 
  html:
    embed-resources: true 
    echo: false
    warning: false
    toc: true
    toc-depth: 3
    toc-location: left
---

```{r}
#| label: setup
#| include: false

library(here)
library(targets)
library(tidyverse)
devtools::load_all()

i_am("analysis/blots.qmd")

tar_config_set(store = here("_targets"))
```

## Experiment

### Drugs in NHLF

```{r}
#| label: exp_ar_sma
#| fig-width: 3.5
#| fig-height: 2.5

tar_read(blot_plot_ar_sma)
```

```{r}
#| label: exp_diclofenac_sma
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_diclofenac_sma)
```

```{r}
#| label: exp_dual_akt
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_akt)
```

```{r}
#| label: exp_dual_pakt
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_pakt)
```

```{r}
#| label: exp_dual_ampk
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_ampk)
```

```{r}
#| label: exp_dual_pampk
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_pampk)
```

```{r}
#| label: exp_dual_cnn1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_cnn1)
```

```{r}
#| label: exp_dual_col1a1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_col1a1)
```

```{r}
#| label: exp_dual_col3a1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_col3a1)
```

```{r}
#| label: exp_dual_erk
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_erk)
```

```{r}
#| label: exp_dual_perk
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_perk)
```

```{r}
#| label: exp_dual_fn1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_fn1)
```

```{r}
#| label: exp_dual_kla
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_kla)
```

```{r}
#| label: exp_dual_mct1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_mct1)
```

```{r}
#| label: exp_dual_mct4
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_mct4)
```

```{r}
#| label: exp_dual_smad3
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_smad3)
```

```{r}
#| label: exp_dual_psmad3
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_psmad3)
```

```{r}
#| label: exp_dual_sma
#| fig-width: 5
#| fig-height: 2.5

tar_read(blot_plot_dual_sma)
```

```{r}
#| label: exp_six_hif1a
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_six_hif1a)
```

### siRNA in NHLF

```{r}
#| label: exp_sirna_mct1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_sirna_mct1)
```

```{r}
#| label: exp_sirna_mct4
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_sirna_mct4)
```

```{r}
#| label: exp_sirna_sma
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_sirna_sma)
```

### TGFβ

```{r}
#| label: exp_untreated_col1a1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_col1a1)
```

```{r}
#| label: exp_untreated_fn1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_fn1)
```

```{r}
#| label: exp_untreated_mct1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_mct1)
```

```{r}
#| label: exp_untreated_mct4
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_mct4)
```

```{r}
#| label: exp_untreated_sma
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_sma)
```

### Lactate Treatment

```{r}
#| label: exp_lactate_col1a1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_col1a1)
```

```{r}
#| label: exp_lactate_fn1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_fn1)
```

```{r}
#| label: exp_lactate_mct1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_mct1)
```

```{r}
#| label: exp_lactate_mct4
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_mct4)
```

```{r}
#| label: exp_lactate_sma
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_sma)
```

### Drugs in Mice

```{r}
#| label: exp_bleo_mct1
#| fig-width: 3.5
#| fig-height: 2.5

tar_read(blot_plot_bleo_mct1)
```

```{r}
#| label: exp_bleo_mct4
#| fig-width: 3.5
#| fig-height: 2.5

tar_read(blot_plot_bleo_mct4)
```

```{r}
#| label: exp_bleo_sma
#| fig-width: 3.5
#| fig-height: 2.5

tar_read(blot_plot_bleo_sma)
```

### IPF Lung Fibroblasts

```{r}
#| label: exp_comp_mct1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_comp_mct1)
```

```{r}
#| label: exp_comp_mct4
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_comp_mct4)
```

```{r}
#| label: exp_comp_sma
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_comp_sma)
```

```{r}
#| label: exp_ipf_lf_col1a1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_ipf_lf_col1a1)
```

```{r}
#| label: exp_ipf_lf_sma
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_ipf_lf_sma)
```

### IPF Lungs

```{r}
#| label: exp_ipf_mct1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_ipf_mct1)
```

```{r}
#| label: exp_ipf_mct4
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_ipf_mct4)
```

```{r}
#| label: exp_ipf_sma
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_ipf_sma)
```

## Protein

### AKT

```{r}
#| label: prot_dual_akt
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_akt)
```

```{r}
#| label: prot_dual_pakt
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_pakt)
```

### AMPK

```{r}
#| label: prot_dual_ampk
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_ampk)
```

```{r}
#| label: prot_dual_pampk
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_pampk)
```

### CNN1

```{r}
#| label: prot_dual_cnn1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_cnn1)
```

### Col1a1

```{r}
#| label: prot_dual_col1a1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_col1a1)
```

```{r}
#| label: prot_ipf_lf_col1a1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_ipf_lf_col1a1)
```

```{r}
#| label: prot_lactate_col1a1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_col1a1)
```

```{r}
#| label: prot_untreated_col1a1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_col1a1)
```

### Col3a1

```{r}
#| label: prot_dual_col3a1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_col3a1)
```

### ERK

```{r}
#| label: prot_dual_erk
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_erk)
```

```{r}
#| label: prot_dual_perk
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_perk)
```

### FN1

```{r}
#| label: prot_dual_fn1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_fn1)
```

```{r}
#| label: prot_lactate_fn1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_fn1)
```

```{r}
#| label: prot_untreated_fn1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_fn1)
```

### HIF-1α

```{r}
#| label: prot_six_hif1a
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_six_hif1a)
```

### Kla

```{r}
#| label: prot_dual_kla
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_kla)
```

### MCT1

```{r}
#| label: prot_bleo_mct1
#| fig-width: 3.5
#| fig-height: 2.5

tar_read(blot_plot_bleo_mct1)
```

```{r}
#| label: prot_comp_mct1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_comp_mct1)
```

```{r}
#| label: prot_dual_mct1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_mct1)
```

```{r}
#| label: prot_ipf_mct1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_ipf_mct1)
```

```{r}
#| label: prot_lactate_mct1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_mct1)
```

```{r}
#| label: prot_sirna_mct1
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_sirna_mct1)
```

```{r}
#| label: prot_untreated_mct1
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_mct1)
```

### MCT4

```{r}
#| label: prot_bleo_mct4
#| fig-width: 3.5
#| fig-height: 2.5

tar_read(blot_plot_bleo_mct4)
```

```{r}
#| label: prot_comp_mct4
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_comp_mct4)
```

```{r}
#| label: prot_dual_mct4
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_mct4)
```

```{r}
#| label: prot_ipf_mct4
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_ipf_mct4)
```

```{r}
#| label: prot_lactate_mct4
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_mct4)
```

```{r}
#| label: prot_sirna_mct4
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_sirna_mct4)
```

```{r}
#| label: prot_untreated_mct4
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_mct4)
```

### Smad3

```{r}
#| label: prot_dual_smad3
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_smad3)
```

```{r}
#| label: prot_dual_psmad3
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_dual_psmad3)
```

### α-SMA

```{r}
#| label: prot_ar_sma
#| fig-width: 3.5
#| fig-height: 2.5

tar_read(blot_plot_ar_sma)
```

```{r}
#| label: prot_bleo_sma
#| fig-width: 3.5
#| fig-height: 2.5

tar_read(blot_plot_bleo_sma)
```

```{r}
#| label: prot_comp_sma
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_comp_sma)
```

```{r}
#| label: prot_diclofenac_sma
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_diclofenac_sma)
```

```{r}
#| label: prot_dual_sma
#| fig-width: 5
#| fig-height: 2.5

tar_read(blot_plot_dual_sma)
```

```{r}
#| label: prot_ipf_lf_sma
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_ipf_lf_sma)
```

```{r}
#| label: prot_ipf_sma
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_ipf_sma)
```

```{r}
#| label: prot_lactate_sma
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_lactate_sma)
```

```{r}
#| label: prot_sirna_sma
#| fig-width: 4
#| fig-height: 2.5

tar_read(blot_plot_sirna_sma)
```

```{r}
#| label: prot_untreated_sma
#| fig-width: 3
#| fig-height: 2.5

tar_read(blot_plot_untreated_sma)
```