### 2024-01-20 analysis/network.qmd
add cosmos analysis

---
title: "Network Analysis"
format: 
  html:
    embed-resources: true 
    echo: false
    warning: false
    toc: true
    toc-depth: 3
    toc-location: left
---

```{r}
#| label: setup
#| include: false

library(here)
library(targets)
library(tidyverse)
devtools::load_all()

i_am("analysis/network.qmd")

tar_config_set(store = here("_targets"))
```

## Build Networks

```{r}
#| label: universe

tt_metab <- tar_read(metab_mcti_tar_tt_tgfb)
all_metab <- tt_metab[["hmdb"]]
sig_metab <- all_metab[tt_metab$adj.P.Val < 0.1]

tt_genes <- tar_read(deg_tgfb) 
all_genes <- tt_genes[["symbol"]]
sig_genes <- all_genes[tt_genes$padj < 0.1]

universe <- 
  hmdb::hmdb_proteins |> 
  filter(hmdb %in% all_metab & protein %in% all_genes)

enriched <- 
  universe |> 
  filter(hmdb %in% sig_metab & protein %in% sig_genes)

edges <- nrow(enriched)
```

## Compare to Random Networks

```{r}
#| label: random-networks

sample_network <- function() {
  metab <- sample(all_metab, length(sig_metab), replace = FALSE)
  gene <- sample(all_genes, length(sig_genes), replace = FALSE)
  hmdb::hmdb_proteins |> 
    filter(hmdb %in% metab & protein %in% gene) |> 
    nrow()
}
edge_list <- replicate(1000, sample_network())

tibble(x = edge_list) |> 
ggplot() +
  geom_histogram(
    aes(x = x), 
    binwidth = 50
  ) + 
  geom_segment(
    x = edges, 
    xend = edges, 
    y = 10, 
    yend = 0, 
    arrow = arrow(length = unit(10, "points")), 
    color = "darkblue", 
    linewidth = 1, 
    lineend = "butt", 
    linejoin = "bevel"
  ) +
  labs(
    title = "Simulated networks", 
    x = NULL, 
    y = "Edge number"
  ) +
  scale_y_continuous(
    expand = expansion(c(0, 0)), 
    limits = nice_limits
  ) +
  theme_plot()
```

## Cluster Network

```{r}
#| label: graph

tidygraph::tbl_graph(edges = select(enriched, hmdb, protein)) |> 
  ggraph::ggraph(layout = "fr") +
  ggraph::geom_edge_link()
```

```{r}
#| label: cluster

clusters <- 
  enriched |> 
  select(red = protein, blue = name) |> 
  as.data.frame() |>
  netZooR::createCondorObject() |> 
  netZooR::condorCluster()
```