### 2024-04-12 manuscript/supplement.qmd

---
title: "Supplementary Material"
subtitle: "Lactate transport inhibition therapeutically reprograms fibroblast 
    metabolism in experimental pulmonary fibrosis"

authors:
- name: David R. Ziehr
  affiliations: 
  - ref: mgh
  - ref: bwh
  - ref: hms
- name: Nathan M. Krah
  affiliations:
  - ref: utah_gen
- name: Kevin Leahy 
  affiliations:
  - ref: bwh
- name: K. Mark Parnell
  affiliations:
  - ref: vettore
- name: Jack Varon
  affiliations: 
  - ref: bwh
  - ref: hms
- name: Rebecca M. Baron
  affiliations:
  - ref: bwh
  - ref: hms
- name: Nancy J. Philp
  affiliations:
  - ref: tj
- name: Lida P. Hariri
  affiliations: 
  - ref: hms
  - ref: mgh_path
- name: Edy Y. Kim
  affiliations:
  - ref: bwh
  - ref: hms
- name: Rachel S. Knipe
  affiliations:
  - ref: mgh
  - ref: hms
- name: Jared Rutter
  affiliations:
  - ref: utah_biochem
  - ref: utah_hhmi
- name: William M. Oldham
  affiliations:
  - ref: bwh
  - ref: hms
  email: woldham@bwh.harvard.edu
  corresponding: true

affiliations:
- id: mgh
  name: Department of Medicine, Massachusetts General Hospital, Boston, MA
- id: bwh
  name: Department of Medicine, Brigham and Women's Hospital and Harvard Medical School, Boston, MA
- id: hms
  name: Department of Medicine, Harvard Medical School, Boston, MA
- id: utah_gen
  name: Department of Human Genetics, University of Utah, Salt Lake City, UT
- id: vettore
  name: Vettore Biosciences, San Francisco, CA
- id: tj
  name: Department of Pathology, Anatomy & Cell Biology, Thomas Jefferson University, Philadelphia, PA
- id: mgh_path
  name: Department of Pathology, Massachusetts General Hospital, Boston, MA
- id: utah_biochem
  name: Department of Biochemistry, University of Utah, Salt Lake City, UT
- id: utah_hhmi
  name: Howard Hughes Medical Institute, University of Utah School of Medicine, Salt Lake City, UT

format:
  docx:
    reference-doc: template.docx
    page-width: 7.5
    output-file: ziehr_2023_mcti_supplement.docx
    fig-align: center

bibliography: 
- library.bib
- pkgs.bib
csl: jci.csl

crossref: 
  fig-title: ""
  title-delim: ""

filters:
- authors-block
---

```{r}
#| label: setup
#| include: false

# ensure refs are up-to-date
targets::tar_load(refs)
targets::tar_load(pkg_citations)

# ensure figures are up-to-date
targets::tar_load(
  c(
    fig01s,
    fig02s, 
    fig03s, 
    fig04s, 
    fig05s, 
    fig06s, 
    fig07s, 
    fig08s, 
    fig09s, 
    fig10s,
    NULL
  )
)
```

\newpage

# Materials and Methods

\newpage

# Figures

![**Fig. S1. Lactate transport inhibition decreases myofibroblast differentiation *in vitro*.** (**A**) Cell count following siRNA-mediated MCT knockdown. (**B**) Small molecule inhibitors of MCT1/2 (AR-C155858, AR) or MCT4 (VB124, VB) decrease TGFβ-stimulated α-SMA expression in normal human lung fibroblasts. (**C**-**D**) Cell counts following pharmacologic inhibition of lactate transporters. Summary data are mean ± SEM (\* p-value-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition). ](figs/SF_01.png)

![**Fig. S2. Lactate transporter inhibition decreases pro-fibrotic gene transcription.** (**A**-**C**) Volcano plots of differentially expressed genes in TGFβ-treated cells compared to control (A), AZD3965-treated cells compared to vehicle (Veh) (B), and VB124-treated cells compared to Veh. Significantly differentially expressed genes are highlighted (adjusted p-value < 0.05), the top 15 up- and down-regulated of which are labeled.](figs/SF_02.png)

\newpage

![**Fig. S3. Lactate transport inhibition alters cellular bioenergetics.** (**A**) Extracellular lactate was determined following TGFβ in the presence of MCT1/2 inhibitor AR-C155858 (AR), MCT4 inhibitor VB124 (VB), or both (N = 3-13 biological replicates, \* p-value < 0.05, *black* compares TGFβ *v.* Ctl, *colored* compares Drug *v.* Veh). (**B**) Extracellular glucose was determined following TGFβ in the presence of MCT1 inhibitor AZD3965 (AZD), VB, or both (N = 5 biological replicates). (**C**) Summary data from Seahorse extracellular flux analysis. (**D**) Cell bioenergetic stress testing results from Seahorse analysis. Glycolytic capacity measures the increase in proton efflux rate following oligomycin treatment. Spare respiratory capacity measures the increase in oxygen consumption rate (OCR) following FCCP treatment compared to basal OCR. Coupling efficiency measures the proportion of basal OCR that contributes to ATP production (*i.e.,* the percentage inhibited by oligomycin). The spare respiratory capacity decreases following MCT inhibition, indicating a larger fraction of total respiratory capacity was utilized in the basal state in cells treated with MCT inhibitors (N = 4 biological replicates, \* p-value < 0.05 compared to TGFβ/Veh). Summary data are mean ± SEM.](figs/SF_03.png)

\newpage

![**Fig. S4. Metabolomic profiling of lung fibroblasts treated with lactate transport inhibitors.** (**A**-**D**) Extracellular (A-B) and intracellular (C-D) metabolomic profile and metabolite set enrichment analyses of lung fibroblasts treated with TGFβ. (**E-H**) Extracellular (E-F) and intracellular (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment. (**I-L**) Extracellular (I-J) and intracellular (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment. Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled. KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by TGFβ treatment or MCT inhibition ordered by normalized enrichment score (NES). Positive NES indicates enrichment in treated cells while negative NES indicates enrichment in Vehicle-treated cells. (**M**) Labeled fraction of intracellular metabolites pyruvate (PYR), lactate (LAC), citrate (CIT), 2-oxoglutarate (2OG), succinate (SUC), and malate (MAL) following treatment with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition). Summary data are mean ± SEM.](figs/SF_04.png)

\newpage

![**Fig. S5. Stable isotope tracing in lung fibroblasts treated with TGFβ.** Lung fibroblasts were cultured with stable isotopes of glucose, lactate, and glutamine during treatment with TGFβ (*columns*). Mass isotope distributions were determined by LC-MS for key metabolites in central carbon metabolic pathways (*rows*). Significant differences in isotopic enrichment are indicated (\* adjusted p-value < 0.05). Summary data are mean ± SEM (N = 4 biological replicates).](figs/SF_05.png)

\newpage

![**Fig. S6. Lactate transport inhibition activates antioxidant defenses.** (**A**) The cellular redox couple NADPH/NADP^+^ was determined by enzymatic cycling assay (N = 4 biological replicates; \* adjusted p-value < 0.05; *black* compares TGFβ to control within a given treatment, *colored* compares the treatment effect to control for a given condition). (**B**) MitoTracker fluorescence was measured as a marker of mitochondrial mass (N = 3 biological replicates; \* adjusted p-value < 0.05 compared to TGFβ-treated cells). (**C**) Proline peak areas determined by LC-MS (N = 5 biological replicates, \* adjusted p-value < 0.05 for the main treatment effect). (**D**) Fraction of the M5 proline isotope following labeling with \[U-^13^C~5~\]-glutamine (N = 4 biological replicates). Summary data are mean ± SEM.](figs/SF_06.png)

\newpage

![**Fig. S7. Mechanisms of lactate signaling.** (**A**) Exogenous lactate (10 mM) does not ameliorate TGFβ-stimulated α-SMA expression. (**B**) Lactate transport inhibition does not attenuate HIF-1α stabilization 6 h after TGFβ treatment. Summary data are mean ± SEM. (N = 4 biological replicates, \* adjusted p-value < 0.05, *black* compares TGFβ *v.* Ctl for a given treatment, *colored* compares treatment *v.* vehicle for the indicated condition).](figs/SF_07.png)

\newpage

![**Fig. S8. Lactate transport inhibition improves weight gain following bleomycin.** Following bleomycin, mice were weighed weekly. Data points show individual mice, summary statistics show the mean ± SEM, \* adjusted p-value-value < 0.05 for the overall treatment effect compared to bleomycin-treated (Bleo) vehicle control.](figs/SF_08.png)

\newpage

![**Fig. S9. Metabolic reprogramming by lactate transport inhibition *in vivo*.** (**A**-**D**) Plasma (A-B) and lung (C-D) metabolomic profile and metabolite set enrichment analyses in bleomycin-treated mice. (**E**-**H**) Plasma (E-F) and lung (G-H) metabolomic profile and metabolite set enrichment analysis of AZD3965 treatment in bleomycin-treated mice. (**I**-**L**) Plasma (I-J) and lung (K-L) metabolomic profile and metabolite set enrichment analysis of VB124 treatment in bleomycin-treated mice. Differentially regulated metabolites are colored (adjusted p-value < 0.1), the top 10 up- and down-regulated of which are labeled. KEGG pathways significantly enriched (adjusted p-value < 0.1) with metabolites differentially regulated by bleomycin treatment or MCT inhibition ordered by normalized enrichment score (NES). Positive NES indicate enrichment in treated mice while negative NES indicate relative enrichment in control or untreated mice.](figs/SF_09.png)

\newpage

![**Fig. S10. VB253 does not attenuate canonical TGFβ signaling.** Upon phosphorylation by the TGFβ receptor, Smad3 relocates to the nucleus. Dose-response profile of VB253 compared to nintedanib (NIN), demonstrates that VB253 does not inhibit canonical Smad3 signaling pathways.](figs/SF_10.png)