### 2024-04-25 README.Rmd
add draft README

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Ziehr.2023.ipf.mcti

<!-- badges: start -->
<!-- badges: end -->

### Author

William Oldham (woldham@bwh.harvard.edu)

### Contents

This research compendium contains all data, code, and text associated with our publication:

> **Lactate transport inhibition therapeutically reprograms fibroblast metabolism in experimental pulmonary fibrosis.** David R. Ziehr, Fei Li, K. Mark Parnell, Nathan M. Krah, Kevin J. Leahy, Christelle Guillermier, Jack Varon, Rebecca M. Baron, Bradley A. Maron, Nancy J. Philp, Lida P. Hariri, Edy Y. Kim, Matthew L. Steinhauser, Rachel S. Knipe, Jared Rutter, and William M. Oldham. *bioRxiv*

### Licenses

| **Text / Figures :** [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)
| **Code :** See the [DESCRIPTION](DESCRIPTION) file.
| **Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/). Attribution requested in reuse.